Formal Grammars as Types in Non-commutative Linear-Non-Linear Type Theory

### 2024-07-10 POPL/paper.tex
comments
addressing some of pedros concerns

% -*- fill-column: 80; -*-
\documentclass[sigconf,anonymous,review,screen,acmsmall]{acmart}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{lipsum}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows, fit}

@@

\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\lto}{\multimap}
\newcommand{\tol}{\mathrel{\rotatebox[origin=c]{180}{$\lto$}}}
\newcommand{\String}{\Sigma^{*}}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Gr}{\mathbf{Gr}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}
\newcommand{\nat}{\mathbb{N}}

@@

\begin{abstract}
  In this paper, we propose a new syntactic formalism for formal grammars: a
  dependent linear/non-linear type theory in which the tensor product is
  non-commutative. The linear types in this calculus may be interpreted as
  grammars, and linear terms as parse transformers. This provides a logical
  characterization of grammar classes based on a substructural logic rather than
  ordinary first-order logic.

  Internally to our calculus, we carry out many classical constructions from
  the theory of formal grammars. Inspired by the common
  structure underlying various ideas in formal language
  theory, our type theory makes these relationships explicit
  and provides a unifying syntax to reason about formal
  grammars.

  We give this type theory a semantics in the category of grammars and prove a
  canonicity theorem showing that every term in a context corresponding to a
  fixed string is equal to a term in a canonical form encoding a parse tree of
  that string. Finally, we discuss the application to the design and
  implementation of correct-by-construction parsers.
\end{abstract}

@@

\paragraph{Contributions.} This paper begins in \cref{sec:categorify} by giving a categorical interpretation of Kleene algebra. However, Kleene algebras are primarily concerned with the regular languages. With this in mind, we further abstract our categorical model from the connectives of regular grammars to those of context-free grammars and beyond. From there, in \cref{sec:tt} we provide a type theory to construct and reason about such formal grammars. In \cref{sec:semantics} we give interpretations of the syntax as grammars and as languages, and prove a canonicity theorem for the type theory. Finally, in \cref{sec:automata} we give a presentation of automata as grammars. With this view in mind, we may encode parsing algorithms and execute them entirely internal to our formalism. Our contributions are then:

\pedro{Something good is coming up with a name to the type theory... This makes it easier to
refer to it throughout the paper. Spitballing some suggestions GTT (grammar type theory --- conflicts with Max's Gradual Type Theory), KTT (Kleene type theory)...}

\begin{itemize}
  \item A universal semantic notion of formal grammar given by the category $\Set^{\String}$
  \item Categorical structures that capture the concepts integral to formal grammar theory, and that are notably modeled by $\Set^{\String}$
  \item A dependent
    linear/non-linear (LNL) type theory to syntactically capture the above categorical structures
  \item An implementation of the language, which is embedded shallowly in Agda
\end{itemize}

\section{A Syntax-Independent Notion of Syntax}

@@

As they are simply sets, formal languages are naturally endowed with a partial
order via subset inclusion. Similarly, formal grammars are naturally endowed
with the structure of a \emph{category}. A morphism $f : G \to H$ between two
grammars is given by a family of functions

@@

for every string $w \in \String$. Intuitively, we read a morphism of grammars as
a \emph{parse transformer}: $f$ translates a $G$-parse of $w$ into an $H$-parse
of $w$. The notion of isomorphism induced by
this categorical structure encodes that two grammars are equivalent if there is
a bijective translation of parses --- precisely Chomsky's notion of \emph{strong equivalence} of
grammars \cite{chom1963}.

\pedro{Written like this, it seems like this correspondence iso-strong
  equivalence is a well-known fact, but this is not an obvious
  observation.}

The above category of grammars can equivalently be described as the functor
category $\Set^{\String}$ --- where the set $\String$ is viewed as a discrete
category. That is, the objects in $\String$ are strings and the only morphisms
are identity morphisms. Such functor categories into $\Set$ --- often called
presheaf categories --- carry a remarkable amount of structure.

The category-theoretic framework can be further used to describe and compare
different notions of formal grammar. For instance, Chomskyan generative
grammars, semi-Thue systems, Montague grammars, and so on are all distinct syntactic
presentations of the same underlying idea of an abstract specification for
parsing. Each of these is an instance of a general \emph{notion of formal
  grammar} --- a category paired with a functor into
$\Set^{\String}$. That is, they may all be interpreted in our category of grammars. In this sense, $\Set^{\String}$ provides a universal setting to reason about parsing.
The slice 2-category over $\Set^{\String}$ then provides the appropriate expressivity to prove equivalences between grammar formalism, such as the equivalence between semi-Thue systems and Chomsky's unrestricted grammars.
\pedro{why?}

@@

\section{Categorifying Formal Grammars}
\label{sec:categorify}
\steven{Idk if this section has the right title, but I wanted to change it if we are adding ``Chomsky categories''}

Kleene algebras are an important tool in the theory of regular
languages. More broadly, through various extensions, they serve as a
theoretical substrate to studying different kinds of formal languages.

We posit that by seeing Kleene algebra through the lens of category
theory, we gain some insight into the type-theoretic essence of formal
grammars. We begin this section by providing a ``simply-typed''
version of a categorification of Kleene algebra, which we call Kleene
category, and then show define a more expressive class of categories,
which we call Chomsky category, that will serve as the categorical
semantics to our type theory.

\subsection{Kleene categories}
A Kleene algebra is a tuple $(A, +, \cdot, (-)^*, 1, 0)$, where $A$ is
a set, $+$ and $\cdot$ are binary operations over $A$, $(-)^*$ is a
function over $A$, and $1$ and $0$ are constants. These structures
satisfy the axioms depicted in Figure~\ref{fig:axioms}.

\begin{lemma}[\cite{day1970construction}]
  Under the assumptions above, the presheaf category $\Set^{\cat{C}}$ is
  monoidal closed.
\end{lemma}

With this in mind, we want to design a categorical semantics that
builds on top of Kleene categories with the goal of extending them
with dependent types and making them capable reasoning about languages
and their parsers. This leads us to the abstract notion of model we
are interested in capturing with our type theory: a \emph{Chomsky
category}.

Further, the syntactic category of the theory in \cref{sec:tt} is manifestly a Chomsky category.
\steven{This is true, right? That's like the entire point of having the syntax}
\pedro{This is likely true, but if we explicitly say so, this warants a proof. I think that
if we don't say anything about the syntactic category, reviewers won't mind.}

\section{Non-commutative Linear-non-linear Type Theory as a Syntax for Formal Grammars}
\label{sec:tt}

@@

\steven{Maybe there is more to say here, or perhaps later, that the only thing
  to gain parsing-wise from a non-dependent non-linear fragment are the things that are
  definable via $G$. Except that these are not interesting, as they are
  necessarily empty}

\steven{Are these paragraphs recounting too much history?}

Inside of an LNL system, linearity may be thought of as an option that users can
choose to deliberately invoke at deliberate points in their developments in an
otherwise intuitionistic system. However, if we are wishing to treat parsers as
linear terms over input strings, the non-linear fragment of an LNL theory does
not really assist in the development of parsers. It is instead the case that
parsers may benefit from a \emph{dependence} on non-linear terms.
\pedro{This last sentence could use an example. Is there a bite-sized, well-known
example we could mention here?}

\steven{Mention separation logic at some point. In particular the separating conjunction and magic wand}

Krishnaswami et al.\ equipped LNL systems with a restricted sense of dependent
types \cite{krishnaswami_integrating_2015}. When defining dependent types in
this setting, dependence on linear terms is disallowed, but either linear or
non-linear terms are allowed to depend on non-linear terms.

The non-linear
subtheory corresponds to Martin-L\"of type theory (MLTT), which can be taken
as a foundation for mathematics. Dependence of linear terms on non-linear terms
of this form effectively allows us to bring any ordinary definitions from math
into scope as indices and then define grammars relative to those indices. That
is all to say, the LNL dependent types in the style of Krishnaswami et al.\
allows us to recover the definition of Aho's indexed grammars \cite{AhoIndexed} internal to our
type theory.
\steven{The above paragraph should maybe be put later. Also flows weirdly}

\pedro{I would argue that both paragraphs above should either be put
  somewhere else or deleted. I think that we should definitely mention
  Neel's LNL system, but as it stands, it feels disconnected from what
  came before and what comes next. A suggestion: the mention of Aho's
  indexed grammars could be the example we use illustrating the
  benefits of dependent grammars.}

@@

We instantiate an LNL theory with dependent types as a deductive
setting to axiomatize the structure of a Chomsky category.

The structural judgments are shown in \cref{fig:structjdg}, the typing
well-formedness in \cref{fig:typewf}, and the intuitionistic typing rules in
\cref{fig:inttyping}. These are all just as they appear in
\cite{krishnaswami_integrating_2015}.

@@

Most of the non-linear semantics is standard, so let us investigate in more
depth the linear semantics. The interpretation of each
connective is exactly what is suggested by the syntax, in
large part to the inspiration taken from this interpretation
when designing the type theory.

@@

That is, strings match $A \lto B$ if when prepended with a parse of $A$ they
complete to parses of $B$. In this manner, the linear function types generalize
Brzozowksi's notion of derivative
\cite{brzozowskiDerivativesRegularExpressions1964}.
Brzozowski initially only gave an accounting of this operation for
generalized regular expressions, but later work from Might et al.\ demonstrates that the same
construction can be generalized to context free grammars
\cite{mightParsingDerivativesFunctional2011}. Here, via the linear function
types the same notion of derivative extends to the grammars of our type theory.

@@

\subsection{Subtheories}
\paragraph{Regular Expressions} Restricting to regular expressions with the
obvious connectives forms a subtheory of our type theory. Perhaps there is
something to say here about how this reduces to Kleene algebra performed inside
of a Kleene cat?

\steven{TODO regular expressions}

\paragraph{$\mu$-Regular Expressions} Regex + fixed point equivalent to context
free \cite{leiss}.

@@

\steven{Every after here has not been touched}

\pedro{After reading this section after the Kleene category one, I
  think that it makes much more sense to switch their order. It's
  going to be easier to justify providing a categorical semantics once
  we have a ``target'' of what we want to interpret.}

\section{Canonicity}
\pedro{It could be cool to come up with a catchy title to this section.
Something like ``Parse Canonicity''}

\pedro{Added the ``initial context'' paragraph below}

Canonicity is an important metatheoretic theorem in the type theory
literature.  It provides insight on the normal forms of terms and,
therefore, on its computational aspects. Frequently, proving
canonicity for boolean types, i.e. every closed term of type bool
reduces to either true or false, is enough to justifying that the type
theory being studied is well-behaved. In our case, however, since we
want to connect our type theory to parsers, we must provide a more
detailed account of canonicity.

@@

@@

Classically, formal language theory is closely related to the study of automata.
In the Chomsky hierarchy, each language class is associated to a class of
automata that serve as recognizers. Internal to the type theory of \cref{sec:tt},
we can characterize these language classes syntactically; moreover, we
demonstrate the equivalence of these language classes to its associated automata class as a
proof term within our logic.

>>>>

% -*- fill-column: 80; -*-
\documentclass[sigconf,anonymous,review,screen,acmsmall]{acmart}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{lipsum}
\usepackage{enumitem}
\usepackage{wrapfig}
\usepackage{fancyvrb}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{automata, positioning, arrows, fit}

@@

\newcommand{\sem}[1]{\llbracket{#1}\rrbracket}
\newcommand{\cat}[1]{\mathbf{#1}}
\newcommand{\lto}{\multimap}
\newcommand{\tol}{\mathrel{\rotatebox[origin=c]{180}{$\lto$}}}
\newcommand{\String}{\Sigma^{*}}
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\SemAct}{\mathbf{SemAct}}
\newcommand{\Gr}{\mathbf{Gr}}
\newcommand{\Grammar}{\mathbf{Gr}}
\newcommand{\Type}{\mathbf{Type}}
\newcommand{\Prop}{\mathbf{Prop}}
\newcommand{\Bool}{\mathbf{Bool}}
\newcommand{\nat}{\mathbb{N}}
\newcommand{\theoryname}{Grammar Type Theory\xspace}
\newcommand{\theoryabbv}{GTT\xspace}

@@

\begin{abstract}
  In this paper we propose \theoryname~(\theoryabbv), a new syntactic formalism for formal grammars: a
  dependent linear/non-linear type theory in which the tensor product is
  non-commutative. The linear types in this calculus may be interpreted as
  grammars, and linear terms as parse transformers. This provides a logical
  characterization of grammar classes based on a substructural logic rather than
  ordinary first-order logic.

  Internal to \theoryname, we carry out many classical constructions from
  the theory of formal grammars. Inspired by the common
  structure underlying various ideas in formal language
  theory, our type theory makes these relationships explicit
  and provides a unifying syntax to reason about formal
  grammars.

  We give a semantics in the category of grammars and prove a
  canonicity theorem showing that every term in a context corresponding to a
  fixed string is equal to a term in a canonical form encoding a parse tree of
  that string. Finally, we discuss the application to the design and
  implementation of correct-by-construction parsers.
\end{abstract}

@@

\paragraph{Contributions.} This paper begins in \cref{sec:synindnotion} by
giving a universal semantic characterization of formal grammar.
From there, in \cref{sec:tt} we provide the syntax of \theoryname, a type thoery
that axiomatizes the underlying structures needed to conduct formal grammar
theory. Then in \cref{sec:categorify}, we characterize the abstract categorical
models of our syntax. Afterwards, we demonstrate several models of \theoryabbv,
include one to prove a metatheoretic canonicity theorem. Finally, in
\cref{sec:automata} we give a presentation of automata as grammars. With this
view in mind, we may encode parsing algorithms and execute them entirely
internal to our formalism. Our contributions are then:

\begin{itemize}
  \item A syntax-independent notion of grammar given by the category $\Grammar$
  \item \theoryname: A dependent
    linear-non-linear type theory meant to syntactically capture the latent structures
    in $\Grammar$ that enable parsing
  \item Categorical structures that capture the concepts integral to formal grammar theory, and that are notably modeled by $\Grammar$
  \item An implementation of the language, which is embedded shallowly in Agda
\end{itemize}

\section{A Syntax-Independent Notion of Syntax}
\label{sec:synindnotion}

@@

As they are simply sets, formal languages are naturally endowed with a partial
order via subset inclusion. Similarly, formal grammars naturally coalesce into a
\emph{category}. Define $\Grammar$ to be the category of formal grammars. Whose
objects are grammars, and whose morphisms are functions $f : G \to H$ between two
grammars presented by a family of functions

@@

for every string $w \in \String$. Intuitively, we read a morphism of grammars as
a \emph{parse transformer} --- i.e. $f$ translates a $G$-parse of $w$ into an $H$-parse
of $w$. The induced notion of isomorphism in $\Grammar$ encodes that two grammars are equivalent if there is
a bijective translation of parses. In other words, two grammars are isomorhpic
if they capture the same structural descriptions of strings, which is
precisely Chomsky's notion of \emph{strong equivalence} of
grammars \cite{chom1963}. In works like
\cite{yoshinaga2002formal}, this notion of strong equivalence is often treated
like the appropriate choice of ``sameness'' for two grammars. When abstracting over the latent
structure of formal grammars and describing it in the the language of
categories, it is encouraging to see the same definition of equivalence arise
naturally.

Upon inspection of the above definition, $\Grammar$ may equivalently be described as the functor
category $\Set^{\String}$ --- where the set $\String$ is viewed as a
discrete\footnote{Objects in $\String$ comprise the set of all strings, and the
  only morphisms are the identity morphisms.}
category. In fact, in this paper we will take $\Grammar$ to be defined as the
functor category $\Set^{\String}$. Such functor categories into $\Set$ --- often called
presheaf categories --- carry a remarkable amount of structure.

The category-theoretic framework can be further used to describe and compare
different notions of formal grammar. For instance, Chomskyan generative
grammars, semi-Thue systems, Montague grammars, and so on are all distinct syntactic
presentations of the same underlying idea of an abstract specification for
parsing. Each of these is an instance of a general \emph{notion of formal
  grammar} --- a category paired with a functor into
$\Grammar$. That is, they may all be interpreted in our category of grammars. In this sense, $\Grammar$ provides a universal setting to reason about parsing.

@@

\section{\theoryname as a Syntax for Formal Grammars}
\label{sec:tt}

@@

Inside of an LNL system, linearity may be thought of as an option that users can
choose to deliberately invoke at deliberate points in their developments in an
otherwise intuitionistic system. However, if we are wishing to treat parsers as
linear terms over input strings, the non-linear fragment of an LNL theory does
not really assist in the development of parsers. It is instead the case that
parsers may benefit from a \emph{dependence} on non-linear terms.
Through the approach described by Krishaswami et al.\ in
\cite{krishnaswami_integrating_2015},
we may define a restricted form of dependent types. In particular, dependence
on linear terms
is disallowed; however, through dependence of a linear term on a non-linear
index, we recover the definition of Aho's indexed grammars \cite{AhoIndexed}
internal to \theoryabbv.

@@

Below we describe \theoryname~(\theoryabbv), an instance of an LNL theory with dependent types
to serve as a deductive
setting for formal grammar theory. \theoryabbv axiomatizes the necessary structure underlying $\Set^{\String}$ to
specify and construct parsers.

The structural judgments are shown in \cref{fig:structjdg}, the typing
well-formedness in \cref{fig:typewf}, and the intuitionistic typing rules in
\cref{fig:inttyping}. These are all just as they appear in
\cite{krishnaswami_integrating_2015}. The treatment of the non-linear types is
pretty standard, so below we focus on the linear syntax.

@@

Most of the non-linear semantics is standard, so let us investigate in more
depth the linear semantics. The interpretation of each
connective is exactly what is suggested by the syntax, ude in
large part to the inspiration taken from the grammar interpretation
when designing \theoryabbv.

@@

That is, strings match $A \lto B$ if when prepended with a parse of $A$ they
complete to parses of $B$. In this manner, the linear function types generalize
Brzozowksi's notion of derivative
\cite{brzozowskiDerivativesRegularExpressions1964}.
Brzozowski initially only gave an accounting of this operation for
generalized regular expressions, but later work from Might et al.\ demonstrates that the same
construction can be generalized to context free grammars
\cite{mightParsingDerivativesFunctional2011}. Here, via the linear function
types the same notion of derivative extends to the grammars of \theoryabbv.

@@

\subsection{Subtheories}
\paragraph{Regular Expressions} We may realize regular expression as a subtheory
of our language by restricting the type constructors to the
linear unit, characters, $\otimes$, $\oplus$, and Kleene star. Classically, the
languages recognized by these are exactly the regular languages --- the lowest
on the Chomksy hierarchy.

\paragraph{$\mu$-Regular Expressions} Similarly, we may restrict the connectives
to be linear unit, characters, $\otimes$, $\oplus$, and arbitrary recursion.
Instead of regular expressions these correspond to Lei{\ss}'s $\mu$-regular
expressions, which are known to be equivalent to context-free grammars
\cite{leiss}.

\steven{Give this subtheory a name }

\paragraph{Beyond Context-Free Grammars} The previous two subtheories induce as semantics
regular grammars and context-free grammars, respectively; however, by including
the LNL dependent types we may actually express the entirety of the Chomsky
hierarchy. Through use of the LNL dependent types, we may encode indexed
grammars, which are known to be properly between context-free and
context-sensitive grammars within the Chomsky hierarchy \cite{AhoIndexed}. We
will further see in \cref{subsubsec:tm} how to use dependence to encode Turing machines internal to our
calculus, which of course induces a semantics of unrestricted grammars.

@@

\section{Categorifying Formal Grammars}
\label{sec:categorify}

Under the Curry-Howard-Lambek correspondence, the programs of the simply-typed
$\lambda$-calculus may be viewed under any one of the three equivalent views: as
terms in a simple functional language, as proofs in an intuitionisitc
propositional logic, or as morphisms in a Cartesian closed category. In the
same vein, with the syntax provided, we aim to give a
categorical accounting of the semantics of \theoryabbv. Let us begin by paring back
the connectives in the language and only considering the subtheory induced by
regular expressions.

\steven{Added this transition to handle putting the categorical semantics after
  the syntax. I think it sounds weird though}

By giving algebraic laws of manipulation and equivalence to regular expressions,
Kleene algebras are an important tool in the theory of regular
languages. More broadly, through various extensions, they serve as a
theoretical substrate to studying different kinds of formal languages.

We posit that by seeing Kleene algebra through the lens of category
theory, we gain some insight into the type-theoretic essence of formal
grammars. We begin this section by providing a ``simply-typed''
version of a categorification of Kleene algebra, which we call Kleene
category, and then show define a more expressive class of categories,
which we call Chomsky category, that will serve as the categorical
semantics to \theoryabbv.

\subsection{Kleene Categories}
A Kleene algebra is a tuple $(A, +, \cdot, (-)^*, 1, 0)$, where $A$ is
a set, $+$ and $\cdot$ are binary operations over $A$, $(-)^*$ is a
function over $A$, and $1$ and $0$ are constants. These structures
satisfy the axioms depicted in Figure~\ref{fig:axioms}.

\begin{lemma}[Day \cite{day1970construction}]
  Under the assumptions above, the presheaf category $\Set^{\cat{C}}$ is
  monoidal closed.
\end{lemma}

With this in mind, we want to design a categorical semantics that
builds on top of Kleene categories with the goal of extending them
with dependent types and making them capable reasoning about languages
and their parsers. This leads us to the abstract notion of model we
are interested in capturing with \theoryabbv: a \emph{Chomsky
category}.

\steven{Max suggests augmenting the definition of a chomsky category to
  something like two categories $L$ and $U$, and $U$ is cartesian closed I don't
  fully recall the rest.

  My best guess is that you take $L$ a Kleene category with a hierarchy of universes
  two, and you further require that $L$ is $Psh(U)$-enriched, except he
  suggested a further adjective on these presheaves. Perhaps representability?
}

Further, the syntactic category of \theoryname is manifestly a Chomsky category.

\pedro{This is likely true, but if we explicitly say so, this warants a proof. I think that
if we don't say anything about the syntactic category, reviewers won't mind.}

\steven{I agree that we don't want to say anything that opens unnecessary
  questions for proofs we haven't written. However, it seems hard to make the
  case that we have the right categorical model of the syntax if this statement
  isn't true. By restating our definition of Chomksy category, this should be
  obvious or a quick proof}

\section{Nonstandard Models}
\label{sec:nonstandard}

\subsection{Semantic Actions}
Returning to the problem of parsing, the output of a parse usually isn't the
literal parse tree. Rather, the output is the result of some \emph{semantic
  actions} ran on the parse tree, which usually serve to remove some syntactic
details that are unnecessary for later processing.

Given a grammar $G : \String \to \Set$, we define a semantic action to be a set
$X$ with a function $f$ that produces a semantic element from any parse of $G$.

\[
  f : \PiTy w \String {G w \to X}
\]

Further, semantic actions can be arranged into a structured category.
Define $\SemAct$ as the comma
category $\Grammar / \Delta$, where $\Delta : \Set \to \Grammar$ defines a
discrete presheaf. That is, for a set $X$, $\Delta (X)(w) = X$ for all
$w \in \String$. As $\SemAct$ is defined as a comma category, it has a forgetul
functor into $\Grammar$. That is, $\SemAct$ serves as a notion of formal
grammar. Moreover, $\SemAct$ is a model of \theoryabbv.

\steven{It being a notion of formal grammar is distinct from being a model. This
probably warrants a proof}

\steven{TODO semantic actions}

\subsection{Parse Canonicity}
Canonicity is an important metatheoretic theorem in the type theory
literature.  It provides insight on the normal forms of terms and,
therefore, on its computational aspects. Frequently, proving
canonicity for boolean types, i.e. every closed term of type bool
reduces to either true or false, is enough to justify that the type
theory being studied is well-behaved. In our case, however, since we
want to connect \theoryabbv to parsers, we must provide a more
detailed account of canonicity.

@@

\pedro{We should conclude this section by explaining the relevance of the canonicity
theorem. Could also be done before stating the theorem.}

@@

Classically, formal language theory is closely related to the study of automata.
In the Chomsky hierarchy, each language class is associated to a class of
automata that serve as recognizers. Internal to \theoryabbv,
we can characterize these language classes syntactically; moreover, we
demonstrate the equivalence of these language classes to its associated automata class as a
proof term within our logic.
